---
title: "Gene Set Variation Analysis - Between passages"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(dpi=300,fig.width=8, fig.height = 6)
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = TRUE
                      )

## Switch working directory to project dir

knitr::opts_knit$set(root.dir = '..')
```


```{r Load data, include=FALSE}
# Load packages

source("./scripts/0_meta_env_prep.R")

# Construct file paths

# Load data for D5 vs D3 GSVA

fit_GSVA <- readRDS("./output/data_enrichment/GSVA/GSVA_results.RDS")

quant_GSVA <- readRDS("./output/data_enrichment/GSVA/quant_GSVA.RDS")

conditions_interest <- c(
  "P5D3Untreated",
  "P7D3Untreated",
  "P13D3Untreated"
)

```


# Methods

Single sample gene set enrichment analysis using GSVA was used to quantify pathway activity within each sample. Statistical analysis using limma was then used to determine activity changes between hMSC growth phases.

The following databases were used:

-   GO

    -   Biological Processes

    -   Cellular Component

    -   Molecular Function

-   MSigDB

    -   h: Hallmark

    -   c2: Curated

    -   c3: Regulatory targets

-   ReactomePA

-   KEGG

# P7vsP5

## GO

### BP

```{r}

glimmaVolcano(
  x = fit_GSVA$GOBP,
  counts = quant_GSVA$GOBP[, quant_GSVA$GOBP@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOBP[, quant_GSVA$GOBP@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```

### CC

```{r}

glimmaVolcano(
  x = fit_GSVA$GOCC,
  counts = quant_GSVA$GOCC[, quant_GSVA$GOCC@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOCC[, quant_GSVA$GOCC@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```

### MF

```{r}

glimmaVolcano(
  x = fit_GSVA$GOMF,
  counts = quant_GSVA$GOMF[, quant_GSVA$GOMF@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOMF[, quant_GSVA$GOMF@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```

## MSigDB

### h

```{r}

glimmaVolcano(
  x = fit_GSVA$h,
  counts = quant_GSVA$h[, quant_GSVA$h@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$h[, quant_GSVA$h@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```

### c2

```{r}

glimmaVolcano(
  x = fit_GSVA$c2,
  counts = quant_GSVA$c2[, quant_GSVA$c2@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$c2[, quant_GSVA$c2@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```

### c3

```{r}

glimmaVolcano(
  x = fit_GSVA$c3,
  counts = quant_GSVA$c3[, quant_GSVA$c3@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$c3[, quant_GSVA$c3@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```

## Reactome

```{r}

glimmaVolcano(
  x = fit_GSVA$Reactome,
  counts = quant_GSVA$Reactome[, quant_GSVA$Reactome@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$Reactome[, quant_GSVA$Reactome@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```

## KEGG

```{r}

glimmaVolcano(
  x = fit_GSVA$KEGG,
  counts = quant_GSVA$KEGG[, quant_GSVA$KEGG@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$KEGG[, quant_GSVA$KEGG@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```

## GCN

```{r}

glimmaVolcano(
  x = fit_GSVA$GCN,
  counts = quant_GSVA$GCN[, quant_GSVA$GCN@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GCN[, quant_GSVA$GCN@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 13
)

```


# P13vsP7

## GO

### BP

```{r}

glimmaVolcano(
  x = fit_GSVA$GOBP,
  counts = quant_GSVA$GOBP[, quant_GSVA$GOBP@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOBP[, quant_GSVA$GOBP@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```

### CC

```{r}

glimmaVolcano(
  x = fit_GSVA$GOCC,
  counts = quant_GSVA$GOCC[, quant_GSVA$GOCC@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOCC[, quant_GSVA$GOCC@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```

### MF

```{r}

glimmaVolcano(
  x = fit_GSVA$GOMF,
  counts = quant_GSVA$GOMF[, quant_GSVA$GOMF@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOMF[, quant_GSVA$GOMF@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```

## MSigDB

### h

```{r}

glimmaVolcano(
  x = fit_GSVA$h,
  counts = quant_GSVA$h[, quant_GSVA$h@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$h[, quant_GSVA$h@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```

### c2

```{r}

glimmaVolcano(
  x = fit_GSVA$c2,
  counts = quant_GSVA$c2[, quant_GSVA$c2@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$c2[, quant_GSVA$c2@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```

### c3

```{r}

glimmaVolcano(
  x = fit_GSVA$c3,
  counts = quant_GSVA$c3[, quant_GSVA$c3@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$c3[, quant_GSVA$c3@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```

## Reactome

```{r}

glimmaVolcano(
  x = fit_GSVA$Reactome,
  counts = quant_GSVA$Reactome[, quant_GSVA$Reactome@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$Reactome[, quant_GSVA$Reactome@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```

## KEGG

```{r}

glimmaVolcano(
  x = fit_GSVA$KEGG,
  counts = quant_GSVA$KEGG[, quant_GSVA$KEGG@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$KEGG[, quant_GSVA$KEGG@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```

## GCN

```{r}

glimmaVolcano(
  x = fit_GSVA$GCN,
  counts = quant_GSVA$GCN[, quant_GSVA$GCN@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GCN[, quant_GSVA$GCN@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 14
)

```


# P13vsP5

## GO

### BP

```{r}

glimmaVolcano(
  x = fit_GSVA$GOBP,
  counts = quant_GSVA$GOBP[, quant_GSVA$GOBP@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOBP[, quant_GSVA$GOBP@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

### CC

```{r}

glimmaVolcano(
  x = fit_GSVA$GOCC,
  counts = quant_GSVA$GOCC[, quant_GSVA$GOCC@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOCC[, quant_GSVA$GOCC@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

### MF

```{r}

glimmaVolcano(
  x = fit_GSVA$GOMF,
  counts = quant_GSVA$GOMF[, quant_GSVA$GOMF@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GOMF[, quant_GSVA$GOMF@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

## MSigDB

### h

```{r}

glimmaVolcano(
  x = fit_GSVA$h,
  counts = quant_GSVA$h[, quant_GSVA$h@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$h[, quant_GSVA$h@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

### c2

```{r}

glimmaVolcano(
  x = fit_GSVA$c2,
  counts = quant_GSVA$c2[, quant_GSVA$c2@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$c2[, quant_GSVA$c2@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

### c3

```{r}

glimmaVolcano(
  x = fit_GSVA$c3,
  counts = quant_GSVA$c3[, quant_GSVA$c3@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$c3[, quant_GSVA$c3@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

## Reactome

```{r}

glimmaVolcano(
  x = fit_GSVA$Reactome,
  counts = quant_GSVA$Reactome[, quant_GSVA$Reactome@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$Reactome[, quant_GSVA$Reactome@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

## KEGG

```{r}

glimmaVolcano(
  x = fit_GSVA$KEGG,
  counts = quant_GSVA$KEGG[, quant_GSVA$KEGG@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$KEGG[, quant_GSVA$KEGG@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

## GCN

```{r}

glimmaVolcano(
  x = fit_GSVA$GCN,
  counts = quant_GSVA$GCN[, quant_GSVA$GCN@phenoData@data$condition_ID %in% conditions_interest]@assayData[["exprs"]],
  groups = quant_GSVA$GCN[, quant_GSVA$GCN@phenoData@data$condition_ID %in% conditions_interest]@phenoData@data$condition_ID,
  xlab = "logFC",
  transform.counts = "none",
  coef = 15
)

```

# Volcano3D

```{r}
# Define volcano3D function

draw_volcano3D <- function(fit, quant, highlights = NULL) {
  
fit_sel <- fit
quant_sel <- quant

table_design <- quant_sel@phenoData@data

table_design$condition_ID <- factor(table_design$condition_ID,
          levels = c("P5D3Untreated",
                     "P7D3Untreated",
                     "P13D3Untreated"),
          labels = c("5",
                     "7",
                     "13"))

polar_pvals <- cbind(
  topTable(fit_sel,           number = Inf, sort.by = "none")$P.Value,
  topTable(fit_sel, coef = 13, number = Inf, sort.by = "none")$P.Value,
  topTable(fit_sel, coef = 15, number = Inf, sort.by = "none")$P.Value,
  topTable(fit_sel, coef = 14, number = Inf, sort.by = "none")$P.Value
)

polar_padj <- cbind(
  topTable(fit_sel,           number = Inf, sort.by = "none")$adj.P.Val,
  topTable(fit_sel, coef = 13, number = Inf, sort.by = "none")$adj.P.Val,
  topTable(fit_sel, coef = 15, number = Inf, sort.by = "none")$adj.P.Val,
  topTable(fit_sel, coef = 14, number = Inf, sort.by = "none")$adj.P.Val
)

## Construct volcano3d object

polar_manual <- polar_coords(
  outcome = table_design$condition_ID,
  data = t(quant_sel@assayData$exprs),
  pvals = polar_pvals,
  padj = polar_padj
)


rownames(polar_manual@pvals) <- rownames(quant_sel@assayData$exprs)
colnames(polar_manual@pvals) <- c("ANOVA", "P7vsP5", "P13vsP5", "P13vsP7")
rownames(polar_manual@padj) <- rownames(quant_sel@assayData$exprs)
colnames(polar_manual@padj) <- c("ANOVA", "P7vsP5", "P13vsP5", "P13vsP7")

## Plot

volcano3D(polar_manual,
              axis_angle = 3/6,
              label_size = 18,
          z_axis_title_size  = 18,
          radial_axis_title_size  = 18,
          label_rows = highlights) %>%
  add_animation()
}

```


## GO

### BP

```{r}

draw_volcano3D(fit_GSVA[["GOBP"]],
               quant_GSVA[["GOBP"]])

```

### CC

```{r}

draw_volcano3D(fit_GSVA[["GOCC"]],
               quant_GSVA[["GOCC"]])

```

### MF

```{r}

draw_volcano3D(fit_GSVA[["GOMF"]],
               quant_GSVA[["GOMF"]])

```

## MSigDB

### h

```{r}

draw_volcano3D(fit_GSVA[["h"]],
               quant_GSVA[["h"]],
               highlights = c("HALLMARK_TGF_BETA_SIGNALING", "HALLMARK_TNFA_SIGNALING_VIA_NFKB",
                              "HALLMARK_MYC_TARGETS_V1",
                              "HALLMARK_ESTROGEN_RESPONSE_LATE",
                              "HALLMARK_SPERMATOGENESIS"))

```

### c2

```{r}

draw_volcano3D(fit_GSVA[["c2"]],
               quant_GSVA[["c2"]])

```

### c3

```{r}

draw_volcano3D(fit_GSVA[["c3"]],
               quant_GSVA[["c3"]])

```

## Reactome

```{r}

draw_volcano3D(fit_GSVA[["Reactome"]],
               quant_GSVA[["Reactome"]])

```

## KEGG

```{r}

draw_volcano3D(fit_GSVA[["KEGG"]],
               quant_GSVA[["KEGG"]])

```

## GCN

```{r}

draw_volcano3D(fit_GSVA[["GCN"]],
               quant_GSVA[["GCN"]])

```
