---
title: "Gene Set Enrichment Analysis - D5 vs D3 - Treated vs. Untreated"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:    
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(dpi=300,fig.width=12, fig.height = 12)

## Switch working directory to project dir

knitr::opts_knit$set(root.dir = '..')

## Hide code blocks

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE
                      )

```

```{r Load data, include=FALSE}
# Load packages

source("./scripts/env_prep.R")

# Construct file paths

path_construct <- function(condition) {
  file.path(getwd(),
            'output',
            'data_enrichment',
            'GSEA',
            condition,
            'GSEA_results.RDS')
}

# Load data for D5 vs D3 GSEA

P5_GSEA <- readRDS(path_construct('D5vsD3_P5_TvsUT'))

P7_GSEA <- readRDS(path_construct('D5vsD3_P7_TvsUT'))

P13_GSEA <- readRDS(path_construct('D5vsD3_P13_TvsUT'))


```

```{r Define plot function, include=FALSE}

plot_GSEA <- function(enrichResults,
                      title) {
  plot_title <- ggdraw() +
    draw_label(title,
               fontface = 'bold')
  
  message(str_c("\nDrawing dot plot",
                title,
                sep = " "))
  
  dotplot <-
    plot_grid(dotplot(enrichResults, showCategory = 30),
              ncol = 1,
              nrow = 1)
  
  message(str_c("\nDrawing Gene-Concept Network plot",
                title,
                sep = " "))
  
  cnetplot <-
    plot_grid(
      cnetplot(
        enrichResults,
        showCategory = 8,
        cex_label_category	= 0.6,
        cex_label_gene	= 0.6
      ),
      ncol = 1,
      nrow = 1
    )
  
  
  message(str_c("\nDrawing Enrichment Map",
                title,
                sep = " "))
  
  emapplot <-
    plot_grid(
      emapplot(
        enrichplot::pairwise_termsim(enrichResults),
        cex_label_category	= 0.6
      ),
      ncol = 1,
      nrow = 1
      
    )
  
  message(str_c("\nDrawing UpSet plot",
                title,
                sep = " "))
  
  upsetplot <- plot_grid(enrichplot::upsetplot(enrichResults),
                         ncol = 1,
                         nrow = 1)
  
  message(str_c("\nDrawing ridgeline plot",
                title,
                sep = " "))
  
  ridgeplot <- plot_grid(
    enrichplot::ridgeplot(enrichResults, showCategory = 30),
    ncol = 1,
    nrow = 1
  )
  
  
  message(str_c("\nExporting plots",
                title,
                sep = " "))
  
  list(
    "dotplot" = plot_grid(
      plot_title,
      dotplot,
      ncol = 1,
      rel_heights = c(0.1, 1)
    ),
    "cnetplot" = plot_grid(
      plot_title,
      cnetplot,
      ncol = 1,
      rel_heights = c(0.1, 1)
    ),
    "emapplot" = plot_grid(
      plot_title,
      emapplot,
      ncol = 1,
      rel_heights = c(0.1, 1)
    ),
    "upsetplot" = plot_grid(
      plot_title,
      upsetplot,
      ncol = 1,
      rel_heights = c(0.1, 1)
    ),
    "ridgeplot" = plot_grid(
      plot_title,
      ridgeplot,
      ncol = 1,
      rel_heights = c(0.1, 1)
    )
  )
}

```

# Methods

After statistical analysis, ordered lists of logFCs between Days 3 and 5 in untreated cells and Days 3 and 5 in treated cells were created. The ordered lists were subjected to GSEA. GSEA was performed against the following datasets:

-   GO

    -   Biological Processes

    -   Cellular Component

    -   Molecular Function

-   MSigDB

    -   h: Hallmark

    -   c2: Curated

    -   c3: Regulatory targets

    -   c5: GO

-   ReactomePA

-   KEGG

# Phase A / Passage+5

## GO

### BP

```{r}

plots <- plot_GSEA(P5_GSEA$GOBP,
                   'Phase A - GO Biological Processes')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### CC

```{r}

plots <- plot_GSEA(P5_GSEA$GOCC,
                   'Phase A - GO Cellular Component')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### MF

```{r}

plots <- plot_GSEA(P5_GSEA$GOMF,
                   'Phase A - GO Molecular Function')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)
```

## MSigDB

### h

```{r}

plots <- plot_GSEA(P5_GSEA$MSigDB_h,
                   'Phase A - MSigDB h')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### c2

```{r}

plots <- plot_GSEA(P5_GSEA$MSigDB_c2,
                   'Phase A - MSigDB c2')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### c3

```{r}

plots <- plot_GSEA(P5_GSEA$MSigDB_c3,
                   'Phase A - MSigDB c3')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### c5

```{r}

plots <- plot_GSEA(P5_GSEA$MSigDB_c5,
                   'Phase A - MSigDB c5')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

## ReactomePA

```{r}

plots <- plot_GSEA(P5_GSEA$Reactome,
                   'Phase A - Reactome')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

## KEGG

```{r}

plots <- plot_GSEA(P5_GSEA$KEGG,
                   'Phase A - KEGG')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

# Phase B / Passage+7

## GO

### BP

```{r}

plots <- plot_GSEA(P7_GSEA$GOBP,
                   'Phase A - GO Biological Processes')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### CC

```{r}

plots <- plot_GSEA(P7_GSEA$GOCC,
                   'Phase A - GO Cellular Component')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### MF

```{r}

plots <- plot_GSEA(P7_GSEA$GOMF,
                   'Phase A - GO Molecular Function')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)
```

## MSigDB

### h

```{r}

plots <- plot_GSEA(P7_GSEA$MSigDB_h,
                   'Phase A - MSigDB h')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### c2

```{r}

plots <- plot_GSEA(P7_GSEA$MSigDB_c2,
                   'Phase A - MSigDB c2')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### c3

```{r}

plots <- plot_GSEA(P7_GSEA$MSigDB_c3,
                   'Phase A - MSigDB c3')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### c5

```{r}

plots <- plot_GSEA(P7_GSEA$MSigDB_c5,
                   'Phase A - MSigDB c5')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

## ReactomePA

```{r}

plots <- plot_GSEA(P7_GSEA$Reactome,
                   'Phase A - Reactome')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

## KEGG

```{r}

plots <- plot_GSEA(P7_GSEA$KEGG,
                   'Phase A - KEGG')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

# Phase C / Passage+13

## GO

### BP

```{r}

plots <- plot_GSEA(P13_GSEA$GOBP,
                   'Phase A - GO Biological Processes')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### CC

```{r}

plots <- plot_GSEA(P13_GSEA$GOCC,
                   'Phase A - GO Cellular Component')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### MF

```{r}

plots <- plot_GSEA(P13_GSEA$GOMF,
                   'Phase A - GO Molecular Function')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)
```

## MSigDB

### h

```{r}

plots <- plot_GSEA(P13_GSEA$MSigDB_h,
                   'Phase A - MSigDB h')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### c2

```{r}

plots <- plot_GSEA(P13_GSEA$MSigDB_c2,
                   'Phase A - MSigDB c2')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

### c5

```{r}

plots <- plot_GSEA(P13_GSEA$MSigDB_c5,
                   'Phase A - MSigDB c5')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

## ReactomePA

```{r}

plots <- plot_GSEA(P13_GSEA$Reactome,
                   'Phase A - Reactome')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```

## KEGG

```{r}

plots <- plot_GSEA(P13_GSEA$KEGG,
                   'Phase A - KEGG')

plot(plots$dotplot)

plot(plots$cnetplot)

plot(plots$emapplot)

plot(plots$upsetplot)

plot(plots$ridgeplot)

```
